On the probability of finding a nonphysical solution through shadowing

### 2020-10-23 main.tex

\section{Discussions}
Through a rigorous counter-example, we show that shadowing can lead
to nonphysical solutions.  The conclusion has an immediate implication for
sensitivity analysis methods based on the concept of shadowing \cite{qiqi-lss}\cite{angxiu-lss}\cite{lasagne}.
When the goal is to compute derivatives of ensemble averages, where the ensemble
is distributed according to the physical measure, i.e., the SRB measure, shadowing-based methods can give wrong results.  This has been observed
before and appropriately attributed as ``ergodicity breaking error'' \cite{patrick} \cite{angxiu-error}. In this paper, we construct rigorous analytical examples of this error.
\begin{figure}
    \centering
    \includegraphics[width=0.45\textwidth]{figure/osc_tent_n.png}
    \includegraphics[width=0.45\textwidth]{figure/osc_tent_s.png}
    \caption{The plucked tent map shown at different values of $s$ and $n$, that represent the magnitude and frequency of the applied oscillation to the original tent map ($n=0,s=0$). On the left, $s$ is held constant at 0.5. On the right, $n$ is held constant at 0.}
    \label{fig:map}
\end{figure}
\begin{figure}
    \centering
    \includegraphics[width=0.45\textwidth]{figure/osc_tent_dens_0.5_0.png}
    \includegraphics[width=0.45\textwidth]{figure/osc_tent_dens_0.5_6.png}
    \caption{The effect of increasing $n$ on the probability distribution associated to the plucked tent map. L: The stationary probability distribution of the plucked tent map at $s=0.5, n = 0$. R: The stationary probability distribution of the plucked tent map at $s=0.5, n = 6$.}
    \label{fig:map}
\end{figure}
Even more troubling is the implication for numerical simulations of chaotic
governing equations.  Through the examples in this paper, it is clear that a numerical solution, satisfying
a governing equation perturbed due to numerics, is not expected to shadow a
physical solution of the real governing physics.  What is, if any, the
relation of a numerical solution, such as DNS of turbulent flows, to
the true physics? While this question remains open, we conclude with a final example illustrating 
another problem with assuming the physicality of shadowing solutions: with a small change in a parameter, the stationary probability distribution of a chaotic system \emph{change} drastically, while if shadowing solutions are physical, the change in the statistics must be small when the parameter perturbation is small. Once again, a perturbation of a tent map can be used as an example. 
We consider a plucked tent map as shown in Figure only can shadowing solutions display statistical averages different from physical solutions, but that the difference can be drastic. 
For $10^{-18}$ perturbation (trunction numerically), we get completely different statistics. How can we trust numerical solutions

[Nisha: shadowing perturabtions predict tiny effects. Real effect is a lot more dramatic. Desmos plot.]
give example of numerics that destroys physicalness

>>>>

\section{Discussions}
Through a rigorous counter-example, we show that shadowing can lead
to nonphysical solutions.  The conclusion has an immediate implication for
sensitivity analysis methods based on the concept of shadowing \cite{qiqi-lss}\cite{angxiu-lss}\cite{lasagne}.
When the goal is to compute derivatives of ensemble averages, where the ensemble
is distributed according to the physical measure, i.e., the SRB measure, shadowing-based methods can give wrong results.  This has been observed
before and appropriately attributed as ``ergodicity breaking error'' \cite{patrick} \cite{angxiu-error}. In this paper, we construct rigorous analytical examples of this error.
\begin{figure}
    \centering
    \includegraphics[width=0.45\textwidth]{figure/osc_tent_n.png}
    \includegraphics[width=0.45\textwidth]{figure/osc_tent_s.png}
    \caption{The plucked tent map shown at different values of $s$ and $n$, that represent the magnitude and frequency of the applied oscillation to the original tent map ($n=0,s=0$). On the left, $s$ is held constant at 0.5. On the right, $n$ is held constant at 0.}
    \label{fig:map}
\end{figure}
\begin{figure}
    \centering
    \includegraphics[width=0.45\textwidth]{figure/osc_tent_dens_0.5_0.png}
    \includegraphics[width=0.45\textwidth]{figure/osc_tent_dens_0.5_6.png}
    \caption{The effect of increasing $n$ on the probability distribution associated to the plucked tent map. L: The stationary probability distribution of the plucked tent map at $s=0.5, n = 0$. R: The stationary probability distribution of the plucked tent map at $s=0.5, n = 6$.}
    \label{fig:map}
\end{figure}
Even more troubling is the implication for numerical simulations of chaotic
governing equations.  Through the examples in this paper, it is clear that a numerical solution, satisfying
a governing equation perturbed due to numerics, is not expected to shadow a
physical solution of the real governing physics.  What is, if any, the
relation of a numerical solution, such as DNS of turbulent flows, to
the true physics? For $10^{-18}$ perturbation (due to numerics), 
we may get completely different statistics from a true, physical solution. 
How then can we trust numerical solutions that, even if they are shadowed by
physical solutions, do not share their statistics?

We conclude with a final example illustrating 
another problem with assuming the physicality of shadowing solutions: with a small change in a parameter, the stationary probability distribution of a chaotic system \emph{can} change drastically, while if shadowing solutions are physical, the change in the statistics must be small when the parameter perturbation is small. Once again, a perturbation of a tent map can be used as an example. 
We consider a plucked tent map as shown in Figure only can shadowing solutions display statistical averages different from physical solutions, but that the difference can be drastic.
